---
title: "Day1"
output: html_document
date: "2025-06-23"
---

```{r set up, eval=FALSE}
options(repos = c(
  "CRAN" = "https://cloud.r-project.org",
  "stan-dev" = "https://stan-dev.r-universe.dev",
  "epiforecasts" = "https://epiforecasts.r-universe.dev"
))

install.packages("nfidd", dependencies = TRUE)

library(nfidd)

cmdstanr::install_cmdstan()
```

# Going Further Challenge 1

## Using log normal rather than gamma for the likelihood

### from Probability distributions and parameter estimation

Load lognormal model from the nfidd package

```{r model1, eval=FALSE}
mod <- nfidd_cmdstan_model("lognormal")
mod
```

Get the data

```{r model1data, eval=FALSE}

lognormals <- rlnorm(100, meanlog = 1.57, sdlog = 0.28)

stan_data <- list(
  n = length(lognormals),
  y = lognormals
)
```

Fit and extract samples

```{r model1fit, eval=FALSE}
logn_fit <- nfidd_sample(mod, data = stan_data)
logn_fit

logn_posterior <- as_draws_df(logn_fit$draws())
head(logn_posterior)
```

Generate posterior predictive samples and plot posterior predictive check

```{r model1pred, eval=FALSE}

logn_ppc <- sapply(seq_along(lognormals), function(i) { rlnorm(n = length(lognormals), meanlog = logn_posterior$meanlog[i],
         sdlog = logn_posterior$sdlog[i]) })

ppc_dens_overlay(y = lognormals, yrep = logn_ppc)

```

# Going Further Challenge 2

## Using gamma rather than log-normal

### From delay-distributions

```{r model2, eval=FALSE}

mod <- nfidd_cmdstan_model("gamma")
mod

```

Specify the time from onset to hospitalisation

```{r model2data, eval=FALSE}
df_onset_to_hosp <- df |>
  mutate(onset_to_hosp = hosp_time - onset_time) |>
  # exclude infections that didn't result in hospitalisation
  drop_na(onset_to_hosp)
```

Use the data to sample from the model posterior

```{r model2fit, eval=FALSE}

res <- nfidd_sample(
  mod,
  data = list(
    N = nrow(df_onset_to_hosp),
    y = df_onset_to_hosp$onset_to_hosp
  )
)

res$summary() #lp_ values are negative
res #ess not great
```

# Simulating Censoring with Floor Division

## Example:

Daily censoring (interval = 1)

```{r censoring1, eval=FALSE}
times <- c(0.2, 0.8, 1.3, 1.9, 2.1, 2.7)
censored <- floor(times / 1) * 1
# Result: [0, 0, 1, 1, 2, 2] 
# demonstrates why the rounding works because our real data is 'times' but we observed 'results'
```

Weekly censoring (interval = 7)

```{r censoring7, eval=FALSE}
days <- c(2, 8, 12, 15, 20)
weekly_censored <- floor(days / 7) * 7
#Result: [0, 7, 7, 14, 14]
```

## For censoring reporting delays:

Simulate continuous delays

```{r, eval=FALSE}
## Simulate continuous delays
true_delays <- rgamma(100, shape = 2, rate = 0.5)
## Apply daily censoring
daily_censored <- floor(true_delays / 1) * 1
## Apply weekly censoring
weekly_censored <- floor(true_delays / 7) * 7
```

Generate example data

```{r, eval=FALSE}
set.seed(123)
continuous_times <- runif(50, 0, 10)
daily_censored <- floor(continuous_times / 1) * 1
weekly_censored <- floor(continuous_times / 7) * 7
```

Plot

```{r, eval=FALSE}
data.frame(
  continuous = continuous_times,
  daily = daily_censored,
  weekly = weekly_censored
) |>
  tidyr::pivot_longer(everything(), names_to = "type", values_to = "time") |>
  ggplot(aes(x = time, fill = type)) +
  geom_histogram(alpha = 0.7, position = "identity", bins = 20) +
  facet_wrap(~type, ncol = 1) +
  labs(title = "Effect of Different Censoring Intervals")
```

##Example of weekly censored data Use the floor() function to round down to integers

```{r, eval=FALSE}

df_dates_week <- df |>
  mutate(
    infection_time = floor(infection_time/7)*7,
    onset_time = floor(onset_time/7)*7,
    hosp_time = floor(hosp_time/7)*7
  )
head(df_dates_week)

df_dates_week <- df_dates_week |>
  mutate(
    incubation_period = onset_time - infection_time,
    onset_to_hosp = hosp_time - onset_time
  )
summary(df$onset_time - df$infection_time)
summary(df_dates_week$incubation_period)
```
